Modeling Environmental Stochasticity

### 2022-04-21 notes/environmental stochasticity.Rmd
fix random intercept + slope to include intercept.

```{r}
tic()
m2 <- bam(
  surv ~ s(log_size_prev, bs = "cr", k = 20) +
    s(year_fct, log_size_prev, bs = "re"),
  family = binomial,
    data = df,
    method = "fREML"
)
toc()
draw(m2)
df <- df %>% mutate(fitted2 = predict(m2, type = "response"))
```

@@

The last one fits the data best, but also takes the longest to fit.
m2 is not far off with a dAIC of 5.

@@

>>>>

```{r}
tic()
m2 <- bam(
  surv ~ s(log_size_prev, bs = "cr", k = 20) +
    s(year_fct, bs = "re") +
    s(year_fct, log_size_prev, bs = "re"),
  family = binomial,
    data = df,
    method = "fREML"
)
toc()
draw(m2)
df <- df %>% mutate(fitted2 = predict(m2, type = "response"))
```

@@

The last one fits the data best, but also takes the longest to fit.
m2 is not far off with a dAIC of 6.8.

@@

With newdata to better reflect what the IPM is doing under the hood:

```{r}
newdata <- expand.grid(
  log_size_prev = seq(min(df$log_size_prev), max(df$log_size_prev), length.out = 50),
  year_fct = unique(df$year_fct)
)

fitted_newdf<- list(m0 = m0, m1 = m1, m2 = m2, m3 = m3) %>% 
  map_dfc(~predict(.x, newdata = newdata, type = "response")) %>% 
  bind_cols(newdata)

p0 <- ggplot(fitted_newdf, aes(x = log_size_prev)) +
  geom_line(aes(y = m0)) +
  labs(title = "no random effect")
p1 <- ggplot(fitted_newdf, aes(x = log_size_prev)) +
  geom_line(aes(y = m1, color = year_fct)) +
  labs(title = "random intercept")
p2 <- ggplot(fitted_newdf, aes(x = log_size_prev)) +
  geom_line(aes(y = m2, color = year_fct)) +
  labs(title = "random intercept and slope")
p3 <- ggplot(fitted_newdf, aes(x = log_size_prev)) +
  geom_line(aes(y = m3, color = year_fct)) +
  labs(title = "random shape")
```

```{r}
p0 + p1 + p2 + p3 &
  labs(y = ".fitted") &
  theme(legend.position = "none") &
  plot_annotation(title = "survival probability") & ylim(0,1)
```
